---
title: "Finding probabilities"
author: "SMWadgymar, KELevasseur"
date: '2024-02-08'
output: html_document
---

A probability distribution is a mathematical description of the possible outcomes of a random variable. There are many types of distributions that have been characterized mathematically and are useful to us when analyzing data.

The binomial distribution is the distribution that describes the frequency of outcomes for each set of trials (or draws or flips of a coin) where there are exactly two possible outcomes for each trial (e.g. heads or tails, survives or dies, yes or no, etc.). The notation used to describe the binomial distribution, B(n,p), includes two parameters: n is the size (# flips) and p is the probability that each is heads (for a fair coin, this is 0.5). A variable is binomially distributed if (1) the trials are independent, (2) the number of trials/flips is fixed, (3) each trial/flip outcome can be classified as a success or failure, and (4) the probability of success is the same for each trial.

R has many built-in functions that allow us to draw samples from well-characterized distributions. We’ll make use of three of those functions for the binomial distribution here: ‘rbinom’, ‘dbinom’, and ‘pbinom’. Importantly, we will make use of these functions to examine probabilities in two ways. First, we’ll run simulations. Simulations are digital experiments. We use them to conduct large, replicated experiments that would not be possible in real life, and as in real life, we want our outcomes to be the product of random chance. The goal is to run enough simulations that the probabilities that we estimate are likely very close approximations of the true probabilities (which are often unknown). For our second approach, we will calculate exact probabilities. We can do this because we are drawing our samples from a well-characterized distribution where the exact probabilities are known.




####Getting to know the 'rbinom' function: simulating a random draw/flip from a binomial distribution.


1) Let's simulate a random draw from a binomial distribution. The rbinom function has three arguments: the number of random draws/flips, the number of "coins" being flipped on each draw, and the probability of a heads as the outcome. Heads will be assigned a value of 1 and tails a value of 0.

```{r}
rbinom(1, 1, 0.5)
```

Q: What does the output here mean?

A: 


2) How would you adjust the code above to simulate ten flips of a single coin?

```{r}
rbinom(???, ???, 0.5)
```

Q: What does the output here mean?

A: 


3) How about one draw of ten coins?

```{r}
rbinom(???, ???, 0.5)
```

Q: What does the output here mean?

A: 


4) How about ten draws with each having 10 coins?

```{r}

```

Q: What does the output here mean?

A: 


5) What would the outcome of 10 trials of 10 coins each be if you were using a biased coin that gave you an 80% chance of producing heads instead of a 50% chance?

```{r}

```

Q: How does the output of this line of code differ from the output you saw in the previous example that used a fair coin?  Did you expect these differences?

A: 












####Let's find probabilities through simulation:

6) Simulate 100000 draws each with 10 fair coins and save the outcome as 'flips'.

```{r}
flips<-rbinom(???, ???, ???)
```

Q: How many values are stored in the variable called flips?  What is your interpretation of each value?

A: 


7) Let's find the fraction of draws where 5 heads occurred.

```{r}
mean(flips==5)
```

Q: Which do you expect will be lower: the probability that 3 out of 10 flips will be heads or the probability that 5 out of 10 flips will be heads? Why?  

A: 


8) Next, calculate the probability that 3 out of 10 flips will be heads. This reflects the probability that this outcome will occur, and is called the density of this distribution at the point where x=3.

```{r}

```

Q: How does the probability that 3/10 flips will be heads compare to the probability that 5/10 flips will be heads?

A:


9) Let's find the cumulative probability that heads will occur 4 or fewer times during each draw. This is referred to as the cumulative probability because it sums up the probability that the number of heads is 0, 1, 2, 3, or 4 out of 10 flips.

```{r}
mean(flips<=4)
```

Q: How would you adjust this line of code to estimate the cumulative probability that heads will occur 6 or more times?

A:
```{r}

```













####Now we'll practice finding exact probabilities:

10) The dbinom function estimates the exact probability density at a given point. The arguments here are the density being estimated, the number of coins, and the probability of producing a head. Note that the order and type of arguments used here do not match the order and type of arguments that you used for the 'rbinom' function.

```{r}
dbinom(5, 10, 0.5)
```

Q: How does this command compare to the steps you took for question 6 and 7 above? Are the estimates comparable? Why or why not?

A: 


11) Now let's estimate the exact cumulative probability density for a given range. The arguments here are the density being estimated, the number of coins, and the probability of producing a head. Note that the order and type of arguments used here do not match the order and type of arguments that you used for the 'rbinom' function.

```{r}
pbinom(4, 10, 0.5)
```

Q: How does this command compare to the steps you took for question 9? Are the estimates comparable? Why or why not?

A: 


