---
title: "Probability exercises"
author: "SMWadgymar, KELevasseur"
date: '2022-08-16'
output: html_document
---

####Exercises

1) Simulating coin flips: Generate 10 separate random flips with probability 0.7 of producing heads.

```{r}

```

Q: What kind of values do you see and what do they represent?

A: 


2) Simulating many draws from a binomial distribution: Generate 100 occurrences of flipping 10 coins, each with 70% probability of producing heads. Produce a table of the probability distribution that you generate (hint: name your table "flips" to make it easier, and your table should report probabilities or proportions, not counts).

```{r}

```

Q: What kind of values do you produce and what do they represent? Plot a histogram and describe the shape that your probability distribution takes.

A: 

```{r}
hist(???, freq=F)
```

Q: What does the freq=F argument mean? Hint: make the same graph on a second line of code, but remove the freq=F part.

A:


3) Calculate the exact probability that 2 heads will arise from 10 coin flips with a 70% probability of coming up tails.  

```{r}

```

Q: Compare your answer with a simulation of 10,000 trials. Do the two approaches yield similar results?

A: 

```{r}

```




4) Calculate the cumulative probability that at least five coins out of 10 are heads with a 30% probability of coming up heads.  

```{r}

```

Q: Compare your answer with a simulation of 10,000 trials.  Do the two approaches yield similar results?

A: 
```{r}

```




5) Repeat the simulation you ran in exercise (3) where you estimated the exact probability that two heads will occur with 10, 100, 1,000, 10,000, and 100,000 trials. Produce a plot depicting the number of trails on the x-axis and the associated probabilities you calculated on the y-axis.  

Hint: The c() function lets you combine values into a vector or list. You will need to make a vector of probabilities for the y axis and a vector of your sample sizes for the x axis using c(#, #, #, #, #). Remember that you can assign your vectors names to make coding easier.

```{r}

```


Produce a second plot but this time log-transform the number of trials. Make sure to adjust your axis labels appropriately.  

```{r}

```


Q: Which simulation yields a result most similar to the exact probability? What pattern do you see in your first graph? What does your second graph reveal that your first graph did not?

A: 



